---
title: "Police Fitness README"
author: "Derek Anderson"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: gfm
---

```{r setup, include=FALSE}
library(knitr)
library(fs)
```

# README

## Introduction

My name is [Derek Anderson](https://github.com/dbanderson1). I am graduate student in Kinesiology at Auburn University. This repository
contains a reproducible pipeline for generating police wellness reports which is also my final project for [PLPA 6820 (Reproducible
Science)](https://bulletin.auburn.edu/coursesofinstruction/plpa/) taught by [Zach Noel](https://agriculture.auburn.edu/about/directory/faculty/zachary-noel/).

## Overview

The scripts found in this project may be beneficial to individuals who are conducting wellness initiatives and work with military, police, fire, or emergency medical services. The overall project is split into two scripts.

### Script #1 - catagorize_data.R

The first script, [catagorize_data.R](https://github.com/dbanderson1/police_fitness/blob/main/categorize_data.R), takes raw fitness data for each participant and places them into ordinal categories. This script is written entirely in base R and requires no additional libraries. The output of this script is later used to generate individual wellness reports. It could also be used to provide summary reports to leadership.

Here you’ll find R script to categorize:

- Body Mass Index (BMI)
- Body Fat Percentages (BF%)
- Sit and Reach
- YMCA Bench Press
- VO2 Max
- Grip Strength
- Vertical Jump
- Forward Lean Balance Test
- Functional Movement Screen (FMS)

The raw data for your script should be placed in the /input folder. It should have the following file structure:

```{r}
colnames(read.csv("input/mock_data.csv"))
```

The R script, when ran, will categorize performance for the majority of the variables into a file called mock_data_graded.csv which is stored in the /output folder.

### Script #2 - wellness_report_ind.Rmd

The second script, [wellness_report_ind.Rmd](https://github.com/dbanderson1/police_fitness/blob/main/wellness_report_ind.Rmd), will take the outputs from the first script and generate individualized wellness reports. The reports will give officers an overview of their physical fitness and is broken into four sections. 

1. Body Composition
2. Physical Performance
3. Flexibility and Mobility
4. Comparison to the Force - for this section it is essential that you have completed all of your population's physical testing. It will show them where they fall on a histogram of their peers and rank them into a percentile.

[This](https://github.com/dbanderson1/police_fitness/blob/main/output/wellness_report_ind.pdf) is a sample what the reports look like.

## Instructions on Use

### Step 1. Clone the Repository

To clone this repository you can copy the following code into your R studio console:

```{bash clone, eval=FALSE}
git clone https://github.com/dbanderson1/police_fitness.git
```

### Step 2. Place the files

You only need two files to generate these reports:

1. Your raw data files - in the case of this project it is mock_data.csv.
2. Your department logo - which we have called logo.png

Both of those items should be located in the /input folder.

### Step 3. Run the First Script

Run [catagorize_data.R](https://github.com/dbanderson1/police_fitness/blob/main/categorize_data.R)

### Step 4. Run the Second Script

[wellness_report_ind.Rmd](https://github.com/dbanderson1/police_fitness/blob/main/wellness_report_ind.Rmd)

Note: This script requires only three R packages:

```bash
library(tidyverse) # pulls in ~10+ core packages (dplyr, ggplot2, readr, etc.)
library(knitr) # ties everything together in rendering 
library(tinytex) # helper package that interacts with your LaTeX distribution
```
However, some users may need to install a full distribution of LaTeX, for tinytex to work. The following distributions are recommended:

1. [MiKTeX](https://miktex.org/) for Windows/Linux
2. [MacTeX](https://www.tug.org/mactex/) for macOS.

### Step 5. Confirm functionality

You should be able to delete all files in the output folder and start with your raw ungraded data and be able to generate reports. After deleting, run the first script. Change the code (line 32) in the second script to run a report on a different participant.

```bash
# --------------------------------------------------------------------------
# This is the key line to change to select which report to generate
# --------------------------------------------------------------------------
p <- df[df$Participant_num == "PFS_001", ][1, ]
```

## File Tree

File tree generated using r code and fs():
```{r generate file tree, eval=FALSE}
fs::dir_tree()
```


```{bash file tree, eval=FALSE}
.
├── categorize_data.R # Script One - Categorizes Data for Later Use
├── input # All of your inputs should be placed into this folder
│   ├── logo.png # Your Department's Logo Goes Here !
│   └── mock_data.csv # Your Raw Data Goes Here !
├── output # Here is where all your outputs are generated
│   ├── mock_data_graded.csv # this is your final graded data
│   ├── wellness_report_ind.md # MD version of the reports
│   ├── wellness_report_ind.pdf # PDF version of the reports
│   └── wellness_report_ind_files # this is auto generated and contains the histograms you created
│       └── figure-gfm
│           ├── comparison-plots-1.png
│           ├── comparison-plots-2.png
│           ├── comparison-plots-3.png
│           ├── comparison-plots-4.png
│           ├── comparison-plots-5.png
│           ├── comparison-plots-6.png
│           ├── comparison-plots-7.png
│           ├── comparison-plots-8.png
│           └── comparison-plots-9.png
├── police_fitness.Rproj
├── README.md        # Top Level README
├── README.Rmd
└── wellness_report_ind.Rmd # Script Two - Generates Wellness Reports
```

## Future Directions

This project is the baseline foundation for several additional projects.

1. Development of a third script which will batch process and generate 100+ reports with Participant_num in the file name
2. Overall report using graded data for the Chief of Police
3. Using snippets of this script to develop similar wellness reports for US Army personnel

## Disclaimer

All data contained in this repository are simulated (mock) data and do not represent real individuals or operational police data. These data are provided for illustrative and educational purposes only.

## About

This project was developed using RStudio (version 2026.01.1, Build 403) and R (version 4.5.2). Package dependencies are managed with renv to support reproducibility and ensure consistent functionality across environments and future analyses.
